The Bolocam Galactic Plane Survey: Survey Description and Data Reduction

### 2009-04-20 methods/ms.tex
some more fixes

% -----------------------------------------------------------------------------
% BGPS specific definitions
\def\bgps{BGPS}
\newcommand{\bgpsarea}{150}
\newcommand{\bcamfwhm}{35.6\arcsec}
\newcommand{\ncores}{7628}
\newcommand{\bgpsdepthlow}{20}
\newcommand{\bgpsdepthhigh}{50}
% -----------------------------------------------------------------------------
% Missions
\def\spitzer{{\em Spitzer}}
% -----------------------------------------------------------------------------

@@

We present the Bolocam Galactic Plane Survey (BGPS), a 1.1 mm continuum survey
of \bgpsarea\ square degrees of the Galactic Plane with 130 in the 1st quadrant,
31 in the 2nd, and 6 in the 3rd.  The BGPS represents
the first large area, systematic survey of the Northern Galactic Plane in the
millimeter continuum without pre-selected targets.  The survey has detected
\ncores\ clumps to a limiting non-uniform 1-$\sigma$ noise level in the range \bgpsdepthlow\ and
\bgpsdepthhigh\ m\jyb RMS.  We present maps of 97 contiguous square degrees from
$\lon=-10.5$ to $\lon=86.5$ and $b=\pm0.5$ with a resolution of $\sim$35\arcsec, as
well as regions in IC1396, $\lon=111$ (NGC7538), W3/4/5, and Gem OB1.  The most
striking feature of the maps is the filamentary structure of millimeter emission
throughout the Galaxy.

@@

The Bolocam observations are sensitive to structure between 35\arcsec\ and
about 200\arcsec\ (half the array field of view, see figure
\ref{fig:PCA_Filter}).  An indicative analysis of the spatial filter function
of the iterative mapping scheme is given in \citet{enoch06}, and we present a detailed analysis
of the filter function in our survey in figure \ref{fig:PCA_Filter}.  The data were
sampled onto a uniform 7\farcs2 grid.  Flux density calibrations were performed
using observations of Mars, Uranus, and Neptune, and they are accurate to
$\sim$20\%.  The flux calibration is described in section \ref{sec:FluxCalibration}.

@@

The instrument beam size (PSF) is measured using the planets Uranus
and Neptune which are nearly point sources for Bolocam.
%The semi-diameter of
%Mars over this period ranged from 2.3\arcsec\ to 2.9\arcsec, with
%Neptune $\sim1.1\arcsec$.  Relative to the \bcamfwhm~FWHM Bolocam
%beam, these may be regarded as point sources; the finite size of Mars
%results in a $< 1.5\%$ correction of the observed flux.
The maps of the planets, including all detectors, are fit to a
Gaussian profile.  The beam area is that averaged over all detectors,
and is $1.71\times10^{-6}$ steradians ($35.6$\farcs).  This
allows conversion of the maps, calibrated in Jy beam$^{-1}$, into
surface brightness (MJy sr$^{-1}$).  The uncertainty in the mean beam size
is $.7\%$, leading to an uncertainty in the surface brightness calibration
of \TBD.

@@

Direct comparison with the fluxes listed in Motte et al. requires
consideration of the factor of three difference in the size of the beams.
For example, the the brightest source in the Cyg-X region is DR 21.
While Motte et al. quote a peak flux density per beam of 4.2 \jyb for
the brightest feature in the DR 21 complex (their peak N46) and 17.9 Jy
in a $\sim$ 17\arcsec\ by 23\arcsec box (0.19 x 0.14 pc at an assumed
distance of 1.7 kpc; see Motte et al. Table 1 in their Appendix).
BGPS measures a peak flux of 17.5 \jyb in an effective beam roughly
35\arcsec\ in diameter, indicating that the relative calibration is
similar on the brightest source in common to both surveys.   However,
DR 21 is an extended source.  Fluxes are best compared using sources
that appear compact as observed with the beam of the IRAM 30 meter
telescope.

We have convolved the \citet{motte07} fits files\footnote{{\tt
ftp://cdsarc.u-strasbg.fr/pub/cats/J/A+A/476/1243/fits/}} with a Gaussian
kernel that generates an approximately 35\arcsec\ beam, and compared the
measured fluxes with those from the BGPS.  Additionally, the data values
in Motte et al. are in units of Jankys per pixel, with each pixel subtending
a 4\arcsec\ by 4\arcsec\ area on the sky.  In contrast, the data values
in BGPS are in units of Janskys per beam where the beam is 31\arcsec\
in diameter.   The conversion of units required multiplication of the
convolved images by factor of 8.

The Motte et al. images were convolved with IRAF task
GAUSS using a 3.5 pixel convolution kernel to synthesize maps with an
effective FWHM beam diameter of 35\arcsec ;  the BGPS maps have effective
beam diameter 4\arcsec\ larger than the CSO beam diameter of 31\arcsec\
because of the smearing effects of optical distortion at the edges of the
Bolocam array.  This distortion has not been removed from the Version 1.0
images.  The Bolocam flux calibration is based on the measurement of
primary calibrators using the central pixels of the array which are not
effected by distortion.   The results are shown  in Figure
\ref{fig:CygnusComparison} and tabulated in Table 2.  This figure and
the table shows that BGPS and \citet{motte07} clumps have fluxes
consistent within about 20\% after convolution to a 35\arcsec\ beam and
scaling to Jy/beam.

@@

\FigureTwo{pca_comparison_nolegend_l111}{deconvolution_comparison_nolegend_l111}{The
fractional flux recovered as a function of source size for
well-separated Gaussian sources with FWHM as indicated.
\emph{Left}: 3 (blue), 7 (green), 10 (yellow), 13 (orange), 16 (red), 21
(teal), 26 (lime), 31 (black) PCA components subtracted.  
Note that flux recovery drops to 50\% at 120\arcsec for the 13 PCA
component cleaning used in the released data.  The 3 and 7 PCA cases
show bumps because atmospheric noise is still present at large scales.
The $>$100\% recovery for simulated sources smaller than the deconvolution kernel is a
side-effect; since the deconvolution kernel is smaller than the beam, no real
astrophysical source should ever be smaller than the kernel.
\emph{Right}:
7.2\arcsec (cyan), 14.4\arcsec (black), 21.6\arcsec (red), 31.2\arcsec (green)
deconvolution kernel for 13 PCA components subtracted.  It is clear that the
larger deconvolution kernels generate spurious signal at smaller scales, but
the flux recovery is very close to 1 at the beam size
(35\arcsec)}{fig:PCA_Filter}{1.0}

>>>>

% -----------------------------------------------------------------------------
% BGPS specific definitions
\def\bgps{BGPS}
\newcommand{\bgpsarea}{150}
\newcommand{\bcamfwhm}{31.2\arcsec}
\newcommand{\ncores}{7628}
\newcommand{\bgpsdepthlow}{20}
\newcommand{\bgpsdepthhigh}{50}
% -----------------------------------------------------------------------------
% Missions
\def\spitzer{{\em Spitzer}}
% -----------------------------------------------------------------------------

@@

We present the Bolocam Galactic Plane Survey (BGPS), a 1.1 mm continuum survey
of \bgpsarea\ square degrees of the Galactic Plane with 130 in the 1st quadrant,
31 in the 2nd, and 6 in the 3rd.  The BGPS represents
the first large area, systematic survey of the Northern Galactic Plane in the
millimeter continuum without pre-selected targets.  The survey has detected
\ncores\ clumps to a limiting non-uniform 1-$\sigma$ noise level in the range \bgpsdepthlow\ and
\bgpsdepthhigh\ m\jyb RMS.  We present maps of 97 contiguous square degrees from
$\lon=-10.5$ to $\lon=86.5$ and $b=\pm0.5$ with a resolution of $\sim$\bcamfwhm, as
well as regions in IC1396, $\lon=111$ (NGC7538), W3/4/5, and Gem OB1.  The most
striking feature of the maps is the filamentary structure of millimeter emission
throughout the Galaxy.

@@

The Bolocam observations are sensitive to structure between \bcamfwhm\ and
about 200\arcsec\ (half the array field of view, see figure
\ref{fig:PCA_Filter}).  An indicative analysis of the spatial filter function
of the iterative mapping scheme is given in \citet{enoch06}, and we present a detailed analysis
of the filter function in our survey in figure \ref{fig:PCA_Filter}.  The data were
sampled onto a uniform 7\farcs2 grid.  Flux density calibrations were performed
using observations of Mars, Uranus, and Neptune, and they are accurate to
$\sim$20\%.  The flux calibration is described in section \ref{sec:FluxCalibration}.

@@

The instrument beam size (PSF) is measured using the planets Uranus
and Neptune which are nearly point sources for Bolocam.
%The semi-diameter of
%Mars over this period ranged from 2.3\arcsec\ to 2.9\arcsec, with
%Neptune $\sim1.1\arcsec$.  Relative to the \bcamfwhm~FWHM Bolocam
%beam, these may be regarded as point sources; the finite size of Mars
%results in a $< 1.5\%$ correction of the observed flux.
The maps of the planets, including all detectors, are fit to a
Gaussian profile.  The beam area is that averaged over all detectors,
and is $1.71\times10^{-6}$ steradians (\bcamfwhm).  This
allows conversion of the maps, calibrated in Jy beam$^{-1}$, into
surface brightness (MJy sr$^{-1}$).  The uncertainty in the mean beam size
is $.7\%$, leading to an uncertainty in the surface brightness calibration
of \TBD.

@@

Direct comparison with the fluxes listed in Motte et al. requires
consideration of the factor of three difference in the size of the beams.
For example, the the brightest source in the Cyg-X region is DR 21.
While Motte et al. quote a peak flux density per beam of 4.2 \jyb for
the brightest feature in the DR 21 complex (their peak N46) and 17.9 Jy
in a $\sim$ 17\arcsec\ by 23\arcsec box (0.19 x 0.14 pc at an assumed
distance of 1.7 kpc; see Motte et al. Table 1 in their Appendix).
BGPS measures a peak flux of 17.5 \jyb in an effective beam roughly
\bcamfwhm\ in diameter, indicating that the relative calibration is
similar on the brightest source in common to both surveys.   However,
DR 21 is an extended source.  Fluxes are best compared using sources
that appear compact as observed with the beam of the IRAM 30 meter
telescope.

We have convolved the \citet{motte07} fits files\footnote{{\tt
ftp://cdsarc.u-strasbg.fr/pub/cats/J/A+A/476/1243/fits/}} with a Gaussian
kernel that generates an approximately \bcamfwhm\ beam, and compared the
measured fluxes with those from the BGPS.  Additionally, the data values
in Motte et al. are in units of Jankys per pixel, with each pixel subtending
a 4\arcsec\ by 4\arcsec\ area on the sky.  In contrast, the data values
in BGPS are in units of Janskys per beam where the beam is 31\arcsec\
in diameter.   The conversion of units required multiplication of the
convolved images by factor of 8.

The Motte et al. images were convolved with IRAF task
GAUSS using a 3.5 pixel convolution kernel to synthesize maps with an
effective FWHM beam diameter of \bcamfwhm ;  the BGPS maps have effective
beam diameter 4\arcsec\ larger than the CSO beam diameter of 31\arcsec\
because of the smearing effects of optical distortion at the edges of the
Bolocam array.  This distortion has not been removed from the Version 1.0
images.  The Bolocam flux calibration is based on the measurement of
primary calibrators using the central pixels of the array which are not
effected by distortion.   The results are shown  in Figure
\ref{fig:CygnusComparison} and tabulated in Table 2.  This figure and
the table shows that BGPS and \citet{motte07} clumps have fluxes
consistent within about 20\% after convolution to a \bcamfwhm\ beam and
scaling to Jy/beam.

@@

\FigureTwo{pca_comparison_nolegend_l111}{deconvolution_comparison_nolegend_l111}{The
fractional flux recovered as a function of source size for
well-separated Gaussian sources with FWHM as indicated.
\emph{Left}: 3 (blue), 7 (green), 10 (yellow), 13 (orange), 16 (red), 21
(teal), 26 (lime), 31 (black) PCA components subtracted.  
Note that flux recovery drops to 50\% at 120\arcsec for the 13 PCA
component cleaning used in the released data.  The 3 and 7 PCA cases
show bumps because atmospheric noise is still present at large scales.
The $>$100\% recovery for simulated sources smaller than the deconvolution kernel is a
side-effect; since the deconvolution kernel is smaller than the beam, no real
astrophysical source should ever be smaller than the kernel.
\emph{Right}:
7.2\arcsec (cyan), 14.4\arcsec (black), 21.6\arcsec (red), 31.2\arcsec (green)
deconvolution kernel for 13 PCA components subtracted.  It is clear that the
larger deconvolution kernels generate spurious signal at smaller scales, but
the flux recovery is very close to 1 at the beam size
(\bcamfwhm)}{fig:PCA_Filter}{1.0}
